Deep Learning

### 2015-01-31 deepreview.tex
got rid of SVMs, added MLP space distortion figure and caption

\documentclass[10pts]{article}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{natbib}
%\usepackage{float}

@@

Patterns can be classified by first converting each raw input vector into a
vector of feature activations and then collecting evidence for each
different class by computing a differently weighted sum of the feature
activations. If good
features can be designed by hand, it is relatively easy to learn
class-specific weights that optimize
discrimination~\citep{Rosenblatt57}. Widely-used examples of such linear
classification methods are logistic regression and linear support vector
machines~\citep{bishop-book95,Boser92}.

However, even with good hand-crafted features, many interesting tasks
cannot be solved by linear classifiers. A common method to get around this
problem is to transform the feature vector into a higher dimensional vector
through a non-linear mapping, exploiting the fact that collections of
vectors are more likely to be easily separated when they are spread around
a high-dimensional space~\citep{Cover65}. When there are at least
as many dimensions as examples, a linear classifier can always separate the
training examples, and one clever way to obtain a very high-dimensional space is to use a
``kernel function'' that measures the similarity between two vectors.
Using the kernel function, any input vector, ${\bf x}_i$ in the training
set can be used to define a feature whose activation for a new input
vector, ${\bf x}_{\rm new}$ is just the similarity between ${\bf x}_i$ and
${\bf x}_{\rm new}$. Given a kernel function, it is possible to jointly
optimize the choice of training examplars used to define features and the
class-specific weights associated with the features~\citep{Boser92}. The training
examplars used to define features are called ``support vectors'' and Support
Vector Machines~\citep{Boser92} have been very successful, especially for
tasks with a relatively small number of training examples in which features
are individually informative. They do not, however, perform well for tasks
such as object recognition in which the input space is high-dimensional and
the input-output mapping is so complex that computing similarities using a
simple kernel function is ineffective. 
%% THE TEXT BELOW IS NICE BUT TOO MUCH DETAIL ABOUT  SVMS FOR THIS
%% REVIEW.  
%% With a general-purpose feature space
%% and kernel (such as the Gaussian kernel), generalization far from the
%% examples is poor~\citep{Bengio-localfailure-NIPS-2006-small}, which is problematic
%% in high-dimensional spaces. However, kernel machines would perform well if a
%% very good set of features were provided: learning very good
%% features is what deep learning can do.

@@

From the earliest days of pattern recognition~\citep{selfridge,Rosenblatt57}
researchers aimed to replace hand-engineered features by multiple layers of
learned features.  This was eventually achieved by using the
backpropagation algorithm which was discovered independently by several
different groups in the 1970s and 1980s~\citep{Werbos74,Parker85,LeCun85,RHW}.  
Backpropagation is simply an application of
the chain rule of derivatives (left of Figure~\ref{fig:backprop-box}),
which yields an efficient way of computing the gradient of a
smooth objective function with respect to all of the weights 
%% and parameters 
%% FOR THE NOVICE READER INCLUDING BOTH WEIGHTS AND "PARAMTERS" IS CONFUSING.
%% ITS BETTER TO SAY SOMETHING CLEARER EVEN IF IT IS LESS GENERAL OR CORRECT.
in  a layered network (Figure~\ref{fig:backprop-box}). 
The objective function typically measures an error to be minimized: 
the discrepancy between
the actual output of the network on a given training case and the target
output for that case.  The key insight is that the error derivatives with
respect to the {\it ouputs} of a layer can be computed by working
backwards from the error derivatives with respect to its {\em inputs}
to the subsequent layer.
%% THE TRICKY STEP IS COMPUTING dE/dOUT FROM dE/dIN IN THE SUBSEQUENT
%% LAYER.  I THINK WE SHOULD FOCUS ON GETTING OVER THAT IDEA FIRST.
%%  I PLAN TO REDO THE FIGURE TO MAKE IT SIMPLER.  ITS MUCH TOO
%%  CONFUSING FOR A NOVICE AT PRESENT.  WE SHOULD HAVE A SEPARATE
%%  DISCUSSION OF THAT.
Once this backwards recursion has been used to compute
error derivatives for the outputs of all layers, it is easy to compute
error derivatives for the incoming weights of the neurons in each layer 
(Figure~\ref{fig:backprop-box}).

\begin{figure}[htp]
\fbox{
\begin{minipage}{\textwidth}
\centerline{
\includegraphics[width=0.3\linewidth]{chain-rule.pdf}}
\centerline{
\hspace*{.5mm}\fbox{\includegraphics[width=0.49\linewidth]{bpfig_forward.pdf}}
\hspace*{0mm}\fbox{\includegraphics[width=0.49\linewidth]{bpfig_backward.pdf}}
}
\caption{
{\bf Top.} The chain rule of derivatives tells us how
two small effects (that of a small change of $x$ on $y$, and that of $y$ on $z$)
are composed. A small change $\Delta x$ in $x$
gets transformed first into a small change $\Delta y$ in $y$
by getting multiplied by $\frac{\partial y}{\partial x}$
(that is the definition of partial derivative). Similarly,
the change $\Delta y$ creates a change $\Delta z$ in $z$. Substituting one equation
into the other gives the chain rule of derivatives, i.e., how
$\Delta x$ gets turned into $\Delta z$ through multiplication
by the product of $\frac{\partial y}{\partial x}$ and
$\frac{\partial z}{\partial x}$. It also works when $x$, $y$
and $z$ are vectors (and the derivatives are Jacobian matrices). \newline
{\bf Bottom left.} The equations used for computing the forward pass in a neural net
with two hidden layers. At each layer, we first compute the total input,
$z$, to each unit which is a weighted sum of the outputs of the units in
the layer below. Then a non-linear function $f(.)$ is applied to $z$ to get
the output of the unit.  For simplicity, we have omitted bias terms. \newline
{\bf Bottom right.} The equations used for computing the backward pass. At each hidden
layer we compute the error-derivative w.r.t. the output of each unit which
is a weighted sum of the error-derivatives w.r.t. the total inputs to the
units in the layer above. We then convert the error-derivative w.r.t. the
output into the error-derivative w.r.t. the input by multipling it by the
derivative of $f(z)$.  At the output layer, the error-derivative w.r.t. the
output of a unit is computed by differentiating the cost function. This
gives $y_l-t_l$ If the cost function for unit $l$ is $0.5(y_l-t_l)^2$ where
$t_l$ is the target value.
}
\label{fig:backprop-box}
\end{minipage}
}
\end{figure}

>>>>

\documentclass[10pts]{article}
\usepackage{times}
\usepackage{epsfig}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{hyperref}

@@

A neural network has input neurons that can be driven by raw input
signals such as the pixels of an image, forming an {\em input vector},
similarly to how sensory neurons are driven by outside signals like photons or sounds.
Patterns can be classified by first converting each raw input vector into a
vector of feature activations (by analogy to the response of brain neurons
to these raw signals) and then collecting evidence for each
different class by computing a differently weighted sum of the feature
activations. If good
features can be designed by hand, it is relatively easy to learn
class-specific weights that optimize
discrimination~\citep{Rosenblatt57}. Widely-used examples of such {\em linear
classifiers} are {\em logistic regression} and {\em linear support vector
machines}~\citep{bishop-book95,Boser92}.

However, even with good hand-crafted features, many interesting tasks
cannot be solved by linear classifiers. By distorting the space
as in Figure~\ref{fig:space-distortion}, a linear discrimination
becomes possible.
\iffalse
A common method to get around this
problem is to transform the feature vector into a higher dimensional vector
(i.e., corresponding to a much larger set of features)
through a non-linear mapping, exploiting the fact that collections of
vectors are more likely to be easily separated when they are spread around
a high-dimensional space~\citep{Cover65}. When there are at least
as many dimensions as examples, a linear classifier can always separate the
training examples, and one clever way to obtain a very high-dimensional space is to use a
``kernel function'' that measures the similarity between two vectors.
Using the kernel function, any input vector, ${\bf x}_i$ in the training
set can be used to define a feature whose activation for a new input
vector, ${\bf x}_{\rm new}$ is just the similarity between ${\bf x}_i$ and
${\bf x}_{\rm new}$. Given a kernel function, it is possible to jointly
optimize the choice of training examplars used to define features and the
class-specific weights associated with the features~\citep{Boser92}. The training
examplars used to define features are called ``support vectors'' and Support
Vector Machines~\citep{Boser92} have been very successful, especially for
tasks with a relatively small number of training examples in which features
are individually informative. They do not, however, perform well for tasks
such as object recognition in which the input space is high-dimensional and
the input-output mapping is so complex that computing similarities using a
simple kernel function is ineffective. 
%% THE TEXT BELOW IS NICE BUT TOO MUCH DETAIL ABOUT  SVMS FOR THIS
%% REVIEW.  
%% With a general-purpose feature space
%% and kernel (such as the Gaussian kernel), generalization far from the
%% examples is poor~\citep{Bengio-localfailure-NIPS-2006-small}, which is problematic
%% in high-dimensional spaces. However, kernel machines would perform well if a
%% very good set of features were provided: learning very good
%% features is what deep learning can do.
\fi

@@

From the earliest days of pattern recognition~\citep{selfridge,Rosenblatt57}
researchers aimed to replace hand-engineered features by multiple layers of
learned features.  This was eventually achieved by using the
backpropagation algorithm which was discovered independently by several
different groups in the 1970s and 1980s~\citep{Werbos74,Parker85,LeCun85,RHW}.  
Backpropagation is simply an application of
the chain rule of derivatives (Figure~\ref{fig:backprop-box}),
which yields an efficient way of computing the gradient of a
smooth objective function with respect to all of the weights 
%% and parameters 
%% FOR THE NOVICE READER INCLUDING BOTH WEIGHTS AND "PARAMTERS" IS CONFUSING.
%% ITS BETTER TO SAY SOMETHING CLEARER EVEN IF IT IS LESS GENERAL OR CORRECT.
in  a layered network (Figure~\ref{fig:backprop-box}). 
The objective function typically measures an error to be minimized: 
the discrepancy between
the actual output of the network on a given training case and the target
output for that case.  The key insight is that the error derivatives with
respect to the {\it ouputs} of a layer can be computed by working
backwards from the error derivatives with respect to its {\em inputs}
to the subsequent layer.
%% THE TRICKY STEP IS COMPUTING dE/dOUT FROM dE/dIN IN THE SUBSEQUENT
%% LAYER.  I THINK WE SHOULD FOCUS ON GETTING OVER THAT IDEA FIRST.
%%  I PLAN TO REDO THE FIGURE TO MAKE IT SIMPLER.  ITS MUCH TOO
%%  CONFUSING FOR A NOVICE AT PRESENT.  WE SHOULD HAVE A SEPARATE
%%  DISCUSSION OF THAT.
Once this backwards recursion has been used to compute
error derivatives for the outputs of all layers, it is easy to compute
error derivatives for the incoming weights of the neurons in each layer 
(Figure~\ref{fig:backprop-box}).

\begin{figure}[htp]
\fbox{
\begin{minipage}{\textwidth}
\centerline{
\fbox{\includegraphics[width=0.69\linewidth]{netvis-simple-2S.png}}
\fbox{\includegraphics[width=0.28\linewidth]{chain-rule.pdf}}}
\centerline{
\hspace*{.5mm}\fbox{\includegraphics[width=0.49\linewidth]{bpfig_forward.pdf}}
\hspace*{0mm}\fbox{\includegraphics[width=0.49\linewidth]{bpfig_backward.pdf}}
}
\caption{
{\bf Top left.} A multi-layer neural network can distort the
input space to make the classes (whose examples are supposed to
be on the red and blue lines, here) linearly separable. Note how
a regular grid in input space is also transformed. This is an
illustrative example with only two input units, two hidden units
and one output unit, but the networks used for object recognition
or natural language processing contain tens or hundreds of thousand
of units. Reproduced with permission by Chris Olah from~\url{http://colah.github.io/}.
\newline
{\bf Top right.} The chain rule of derivatives tells us how
two small effects (that of a small change of $x$ on $y$, and that of $y$ on $z$)
are composed. A small change $\Delta x$ in $x$
gets transformed first into a small change $\Delta y$ in $y$
by getting multiplied by $\frac{\partial y}{\partial x}$
(that is the definition of partial derivative). Similarly,
the change $\Delta y$ creates a change $\Delta z$ in $z$. Substituting one equation
into the other gives the chain rule of derivatives, i.e., how
$\Delta x$ gets turned into $\Delta z$ through multiplication
by the product of $\frac{\partial y}{\partial x}$ and
$\frac{\partial z}{\partial x}$. It also works when $x$, $y$
and $z$ are vectors (and the derivatives are Jacobian matrices). \newline
{\bf Bottom left.} The equations used for computing the forward pass in a neural net
with two hidden layers. At each layer, we first compute the total input,
$z$, to each unit which is a weighted sum of the outputs of the units in
the layer below. Then a non-linear function $f(.)$ is applied to $z$ to get
the output of the unit.  For simplicity, we have omitted bias terms. \newline
{\bf Bottom right.} The equations used for computing the backward pass. At each hidden
layer we compute the error-derivative w.r.t. the output of each unit which
is a weighted sum of the error-derivatives w.r.t. the total inputs to the
units in the layer above. We then convert the error-derivative w.r.t. the
output into the error-derivative w.r.t. the input by multipling it by the
derivative of $f(z)$.  At the output layer, the error-derivative w.r.t. the
output of a unit is computed by differentiating the cost function. This
gives $y_l-t_l$ If the cost function for unit $l$ is $0.5(y_l-t_l)^2$ where
$t_l$ is the target value.
}
\label{fig:backprop-box}
\end{minipage}
}
\end{figure}
